On the eletrodynamic of moving bodies

### 2019-05-30 Article_AS.Rmd
modif1 intro

Canadian forest is an important part of the country's history, culture, natural environment and economy. But thousands of hectares of canadian forest are harvested each year and an additionnal loss is due to insects, diseases and fires [@DRF2000; @NaturalResourcesCanada2019]. To maintain ressource constant, it is therefore essential to compensate the sampling and the looses by reforesting. The reforestation, when under human control, could ideally be made by using better trees than the former ones, i.e. trees that growth faster, that are more resistant to perturbation and that are of better stuctural quality. To achieve this task, we must take into account the cambial activity that is governed by environmental conditions and genetics [@Dufour2010; @Rossi2012]. Therefore, genetic selection is one of the available tools and is currently becoming a crucial and increasingly necessary for the development of the canadian forest [@Bonfils2004]. It allows to select trees with characters of interest and to cross them or to reproduce them in order to obtain trees with interesting properties for the development of the forests or the quality of the wood for the human use. \ 
Improvement programs by genetic selection are in place for the most important commercial species [@DRF2000; @Mullin2011] including white spruce (Picea glauca (Moench) Voss) which has a significant economic value because of its geographical distribution and abundance [@Eyre1980; @Abrahamson2015]. Spruce wood is valuable as lumber, material for carpentry products and for pulping because of the characteristics of its fibres [@Zhang2008]. In addition, the plasticity of spruce, enable this species to adapt to various types of soil and climatic conditions [@Nienstaedt1990]. All these characteristics make white spruce a species of great interest in North America and consequently it is heavily reforested in various silvicultural systems in Canada with more than 150 million of natural an hybrid seedlings planted yearly accross almost all forested regions which represents a wide range of soils types and a climatic conditions [@Canadian2009]. \

During the first half of the 20th century, most studies about genetic selection were dedicated on increasing wood volume production [@Dhir1976; @nienstaedt1985; @Li1997; @Koubaa2000] but with the increasing demand of high structural quality products and the fear of lossing ressource from natural causes, the aspects of wood quality became as important as wood volume [@Franceschini2012]. Therefore, conifer breeding programs are more and more considering wood quality as selection criteria to improve end-use properties. Wood quality has mainly two components: biologic and economic. The biologic component is the resistance to mortality by vigor and the economic component is the value of the wood for end-use. Several traits were considered to assess wood quality and the density is one of the most widely used indicators and arguably the most important one [@zobel1989; @Saranpaa1994; @Mckendry2002; @Macdonald2002; @Alteyrac2006]. Indeed this trait is important for both components of quality because it is strongly correlated with different mechanical and anatomical wood traits [@zobel1989]. Variation in density within a piece of wood may lead to deformation during transformation and inhomogeneous machining properties. Consequently studies have been set up to understand the genetic sources of phenotypic variations of wood quality [@Corriveau1971; @Li1993; @Hernandez2001; @Beaulieu2003; @Lenz2013; @McLean2016]. For example, future genetic selections in breeding programs for different species of spruce trees in Quebec will consider wood properties such as wood density [@Mullin2011; @Beaulieu2009]. 
With regard to the biological component, in has been demonstrated [@Hietz2013; @osazuwa2014] that trees with a more uniform wood densitiy within the stem are also the more resistant to mortality. For example a uniform density gives a greater mechanical resistance to the wind and therefore a lower probability of mortality during a climatic event such as a storm. With regard to the economical context, uniformity affects the performance of saw timber, the pulp yield and calorific values for biomass applications for bioenergy [@Hernandez2001; @Macdonald2002; @Mckendry2002]. For exemple, a strong interannual variation of the density of the wood leads to a decrease of the mechanical resistance while a relatively uniform density produce wood with better mechanical performance. It is therefore important from both a biological point of view and a end-use point of view to have a wood with a uniform density within the stem. \

In view of the studies cited aboves, we understand that there might be a strong gain at selecting trees that are prone to produce wood of uniform density to get a resistant and higly valuable forest in the next decades. To reach this goal is important to understant where do the variations of density come from. In white spruce, there is a typical low frequency radial pattern of variation from pith to bark [@Moore2011]. First a decrease of density during the first years of growth and then an increase from the beginning of the formation of mature wood [@panshin1980]. In addition, this radial variations there are hight frequency interannual variation [@Nicault2010]. Contrary to the radial variation that is under internal control, the interannual variation of the density is influenced mainly by events events to the tree such as the edaphic and climatic conditions [@Nicault2010] that happen at high frequency. \

Unlike growth that is more controlled by environment than genetic and therefore has low to moderate heritability [@Ivkovich2002a], wood density in white spruce is under strong genetic control with moderate to high heritability [@Corriveau1991; @Yanchuk1993; @Ivkovich2002a; @Ivkovich2002b; @Beaulieu2006] and this makes this trait attractive for genetic improvement. Moreover, Climate is a determining factor for the growth and physiological processes of trees. As a result, climatic conditions strongly affect the properties of wood. In fact, certain climatic factors such as drought can cause stress in trees and cause a variation in wood properties, including density, ring width and conductivity. In addition, long-term growth analyzes have shown that it may increase due to temperature, atmospheric CO2 content and nitrogen deposition [@becker1995; @huang2007; @zaehle2014]. Knowing that density can be negatively correlated with growth in spruce [@zhang1995], it is essential to take climat factor into account.

The timeframe of improved forest management being long [@Zobel1972] it is essential to take into account the current climate change context [@Gray2016]. With the increasing temperature, the trees are expected to growth faster and thus being of lower quality because of lower density [@corriveau1987; @DeBell1994; @Lindstrom1996; @Dutilleul1998; @Bouriaud2005]. It is necessary to anticipate the possible environmental variations because more important frequencies water stress or important precipitations are feared with the climatic changes [@IPCC2007; @pachauri2014]. These environmental variations may lead to an increase or decrease in radial growth and thus cause greater variations in the density of the wood. In addition to selecting trees that are prone to produce wood of uniform density we must also ensure that these properties will persist in time with climatic change. Thus the selection must also considered the stability of the properties in time or, more accuralty, with climate change. furthermore, heritability of density has been proven in several studies, but there is no evidence to suggest that density variability is hereditary. The heritability of this trait could qualify it for the selection criteria for future breeding programs for producing better quality wood.\

In this study we quantified and compared heteregoneity of wood density from trees of known parents in order to figure out if some individus produce more uniform wood than others. Knowing the parents of each tree we tried to characterize the genetic contribution to the homogeneity. By comparing two growth sites we tried to characterize the environmental contribution to homogeity. We ultimalty aim to select and reproduce trees families of interest that are both low-sensitivity to climate change and genetically prone to make homogenous wood. Thus, the most homogeneous families will be identified for reforestation and identify variables to consider in breeding of growing trees with less variable wood.

@@

![Study sites](graph_table/Study.sites.png){width=500px; height=500px}

@@

The model we have developed has made it possible to optimize the density of the wood for each study site. In table 1, we observe that the average value of the parameters differs between Asselin and Saint-Casimir. for the parameters b0 and b1, the average is higher in Saint-Casimir and for parameter b2 it is higher in Asselin.\

|Site                     |Parameter |    Min|   Mean|    Max|
|:------------------------|:---------|------:|------:|------:|
| Asselin (Témiscouata)   |$b_0$     | 257.26| 302.78| 369.78|
|                         |$b_1$     |  -0.93|   0.97|   5.36|
|                         |$b_2$     | -81.83|  17.02|  61.34|
| Saint-Casimir (Portneuf)|$b_0$     | 264.88| 304.40| 353.90|
|                         |$b_1$     |  -1.34|   1.09|   7.04|
|                         |$b_2$     | -50.87|  14.87|  61.34|
Table: Summary table of variability index

@@

![The ring density versus the number of annual rings of the samples for the Asselin site.The bold lines represent the model prediction](graph_table/Density_asselin.png){width=330px; height=330px}

![The ring density versus the number of annual rings of the samples for the Saint-Casimir site.The bold lines represent the model prediction](graph_table/Density_casimir.png){width=330px; height=330px}

Model predictions are similar to data trends across sites (fig3&4). A decrease in density is observed in the first years of growth and then an increase in density from a certain cambial age. Whether it is the data or the predictions of the model, we observe a difference between the Asselin (fig 3) and Saint-Casimir (fig 4) sites. The increase in density after the decrease of the first years of growth is greater in Saint-Casimir than in Asselin. In spite of the interannual variations, the trees have the same general tendency and this one seems well described by the model which adapts to each of the sites.\

@@

|Rank correlation  |  Rho| P-value|
|:-----------------|----:|-------:|
|Variability index | 0.08|    0.43|
|$b_0$             | 0.70|    0.00|
|$b_1$             | 0.80|    0.00|
|$b_2$             | 0.60|    0.00|
Table: Sperman’s Rang correlation

@@

Heritability is higher in each site separately than that of the combined sites for all traits. Wood density is the only character with strong heritability for each of the sites as well as for the combined sites. The annual ring width and the parameter $b_0$ have low heritability. When parameters $b_1$ and $b_2$, their heritability is very low. The uniformity index has almost zero heritability regardless of the study site.

>>>>

### Context
Canadian forest is an important part of the country's history, culture, natural environment and economy. But thousands of hectares of canadian forest are harvested each year and an additionnal loss is due to insects, diseases and fires [@DRF2000; @NaturalResourcesCanada2019]. To maintain ressource constant, it is therefore essential to compensate the sampling and the looses by reforesting. The reforestation, when under human control, could ideally be made by using better trees than the former ones, i.e. trees that growth faster, that are more resistant to perturbation and that are of better stuctural quality. To achieve this task, we must take into account the cambial activity that is governed by environmental conditions and genetics [@Dufour2010; @Rossi2012]. Therefore, genetic selection is one of the available tools and is currently becoming a crucial and increasingly necessary for the development of the canadian forest [@Bonfils2004]. It allows to select trees with characters of interest and to cross them or to reproduce them in order to obtain trees with interesting properties for the development of the forests or the quality of the wood for the human use. \ 
Improvement programs by genetic selection are in place for the most important commercial species [@DRF2000; @Mullin2011] including white spruce (Picea glauca (Moench) Voss) which has a significant economic value because of its geographical distribution and abundance [@Eyre1980; @Abrahamson2015]. Spruce wood is valuable as lumber, material for carpentry products and for pulping because of the characteristics of its fibres [@Zhang2008]. In addition, the plasticity of spruce, enable this species to adapt to various types of soil and climatic conditions [@Nienstaedt1990]. All these characteristics make white spruce a species of great interest in North America and consequently it is heavily reforested in various silvicultural systems in Canada with more than 150 million of natural an hybrid seedlings planted yearly accross almost all forested regions which represents a wide range of soils types and a climatic conditions [@Canadian2009]. \

### Qualité du bois et ses composantes
During the first half of the 20th century, most studies about genetic selection were dedicated on increasing wood volume production [@Dhir1976; @nienstaedt1985; @Li1997; @Koubaa2000] but with the increasing demand of high structural quality products and the fear of lossing ressource from natural causes, the aspects of wood quality became as important as wood volume [@Franceschini2012]. Therefore, conifer breeding programs are more and more considering wood quality as selection criteria to improve end-use properties. Wood quality has mainly two components: biologic and economic. The biologic component is the resistance to mortality by vigor and the economic component is the value of the wood for end-use. With regard to the biological component, in has been demonstrated [@Hietz2013; @osazuwa2014] that trees with a more uniform wood densitiy within the stem are also the more resistant to mortality. For example a uniform density gives a greater mechanical resistance to the wind and therefore a lower probability of mortality during a climatic event such as a storm. With regard to the economical context, several traits were considered to assess wood quality and the density is one of the most widely used indicators and arguably the most important one [@zobel1989; @Saranpaa1994; @Mckendry2002; @Macdonald2002; @Alteyrac2006]. Indeed this trait is important for both components of quality because it is strongly correlated with different mechanical and anatomical wood traits [@zobel1989]. Variation in density within a piece of wood may lead to deformation during transformation and inhomogeneous machining properties. In fact, wood density uniformity affects the performance of saw timber, the pulp yield and calorific values for biomass applications for bioenergy [@Hernandez2001; @Macdonald2002; @Mckendry2002]. For exemple, a strong interannual variation of the density of the wood leads to a decrease of the mechanical resistance while a relatively uniform density produce wood with better mechanical performance.
 It is therefore important from both a biological point of view and a end-use point of view to have a wood with a uniform density within the stem. \

### Ce que l'on sait de l'uniformité du bois et son importance
In view of the studies cited aboves, we understand that there might be a strong gain at selecting trees that are prone to produce wood of uniform density to get a resistant and higly valuable forest in the next decades. Consequently, studies have been set up to understand the genetic sources of phenotypic variations of wood quality [@Corriveau1971; @Li1993; @Hernandez2001; @Beaulieu2003; @Lenz2013; @McLean2016]. For example, future genetic selections in breeding programs for different species of spruce trees in Quebec will consider wood properties such as wood density [@Mullin2011; @Beaulieu2009]. Hense, it is important to understant where do the variations of density come from in order to better integrated them into genetic selections. In white spruce, there is a typical low frequency radial pattern of variation from pith to bark [@Moore2011]. First a decrease of density during the first years of growth and then an increase from the beginning of the formation of mature wood [@panshin1980]. In addition, this radial variations there are hight frequency interannual variation [@Nicault2010]. Contrary to the radial variation that is under internal control, the interannual variation of the density is influenced mainly by events events to the tree such as the edaphic and climatic conditions [@Nicault2010] that happen at high frequency. \

### Contrôle génétique et environnemental de la densité du bois
Unlike growth that is more controlled by environment than genetic and therefore has low to moderate heritability [@Ivkovich2002a], wood density in white spruce is under strong genetic control with moderate to high heritability [@Corriveau1991; @Yanchuk1993; @Ivkovich2002a; @Ivkovich2002b; @Beaulieu2006] and this makes this trait attractive for genetic improvement. Although variation in wood density is strongly controlled by genetics (30% - 60%) in white spruce, some of the variation is explained by other factors such as growth. Knowing that climate is a determining factor for the growth and physiological processes of trees. As a result, climatic conditions affect wood density. In fact, some climatic factors such as drought can cause stress in trees and cause a variation in wood properties, including density, ring width and conductivity. In addition, long-term growth analyzes have shown that it may increase due to temperature, atmospheric CO2 content and nitrogen deposition [@becker1995; @huang2007; @zaehle2014]. Considering that density can be negatively correlated with growth in spruce [@zhang1995], it is essential to take climat factor into account.

### L'importance de la prise en compte du climat et de l'héritabilité des caractères dans les les programmes de selection génétique
The timeframe of improved forest management being long [@Zobel1972] it is important to anticipate future changes in environmental conditions in these programs [@Gray2016]. This will allow the establishment of plantations capable of adapting to the water stress and possible significant rainfall related to climate change because climate predictions predict more variations in the coming years [@IPCC2007; @pachauri2014]. In fact, with the increasing temperature, the trees are expected to growth faster and thus being of lower quality because of lower density [@corriveau1987; @DeBell1994; @Lindstrom1996; @Dutilleul1998; @Bouriaud2005] hence, the importance of integrating these factors into genetic selection programs. However, to selecting trees that are prone to produce wood of uniform density we must also ensure that these properties will persist in time with climatic change. Thus the selection must also considered the stability of the properties in time or, more accurately, with climate change. Furthermore, heritability of density has been proven in several studies, but there is no evidence to suggest that density variability is hereditary. The heritability of this trait could qualify it for the selection criteria for future breeding programs for producing better quality wood.\

### Le but de notre étude et ce que nous avons fait. 
In this study we quantified and compared heteregoneity of wood density from trees of known parents in order to figure out if some individus produce more uniform wood than others. Knowing the parents of each tree we tried to characterize the genetic contribution to the homogeneity. By comparing two growth sites we tried to characterize the environmental contribution to homogeity. We ultimalty aim to select and reproduce trees families of interest that are both low-sensitivity to climate change and genetically prone to make homogenous wood. Thus, the most homogeneous families will be identified for reforestation and identify variables to consider in breeding of growing trees with less variable wood.

@@

![Study sites](graph_table/Study.sites.png){width=400px; height=400px}

@@

The model we have developed has made it possible to optimize the density of the wood for each study site. In table 1, we observe that the average value of the parameters differs between Asselin and Saint-Casimir. for the parameters $b_0$ and $b_1$, the average is higher in Saint-Casimir and for parameter $b_2$ it is higher in Asselin.\

|Site                     |Parameter |    Min|   Mean|    Max|
|:------------------------|:---------|------:|------:|------:|
| Asselin (Témiscouata)   |$b_0$     | 257.26| 302.78| 369.78|
|                         |$b_1$     |  -0.93|   0.97|   5.36|
|                         |$b_2$     | -81.83|  17.02|  61.34|
| Saint-Casimir (Portneuf)|$b_0$     | 264.88| 304.40| 353.90|
|                         |$b_1$     |  -1.34|   1.09|   7.04|
|                         |$b_2$     | -50.87|  14.87|  61.34|
Table: Summary table of parameters

@@

![The ring density versus the number of annual rings of the samples for the Asselin site.The bold lines represent the model prediction](graph_table/Density_asselin.png){width=330px; height=330px} \

![The ring density versus the number of annual rings of the samples for the Saint-Casimir site.The bold lines represent the model prediction](graph_table/Density_casimir.png){width=330px; height=330px} \

Model predictions are similar to data trends across sites (fig 3&4). A decrease in density is observed in the first years of growth and then an increase in density from a certain cambial age. Whether it is the data or the predictions of the model, we observe a difference between the Asselin (fig 3) and Saint-Casimir (fig 4) sites. The increase in density after the decrease of the first years of growth is greater in Saint-Casimir than in Asselin. However, the average density is very similar between the two sites. In spite of the interannual variations, the trees have the same general tendency and this one seems well described by the model which adapts to each of the sites.\

@@

When comparing the index of variability between study sites, there is greater variability in Asselin than in Saint-Casimir. The maximum value of 1155.8 in Asselin, for example, is 746, which represents a huge difference.

|Rank correlation  |  Rho| P-value|
|:-----------------|----:|-------:|
|Variability index | 0.08|    0.43|
|$b_0$             | 0.70|2.2^-16^|
|$b_1$             | 0.80|2.2^-16^|
|$b_2$             | 0.60|2.2^-16^|
Table: Sperman’s Rang correlation

@@

Heritability is higher in each site separately than that of the combined sites for all traits. Wood density is the only character with strong heritability for each of the sites as well as for the combined sites. The annual ring width and the parameter $b_0$ have low heritability. When parameters $b_1$ and $b_2$, their heritability is very low. The uniformity index has almost zero heritability regardless of the study site.\
